### 2021-12-21 slides.tex
Init slides

\documentclass{beamer}
%Information to be included in the title page:
\title{Conformance Checking Using Embeddings and its Applicability in the Internet of Things}
\author{Jan Kruska}
\date{\today}

\begin{document}

\frame{\titlepage}

\begin{frame}
	\frametitle{Table of Contents}
	\tableofcontents
\end{frame}
\section{Conformance Checking}
\begin{frame}
	\frametitle{Conformance Checking}
\end{frame}

\section{Embeddings}
\begin{frame}
	\frametitle{Embeddings}
\end{frame}

\begin{frame}
	\frametitle{\emph{act2vec} \& \emph{trace2vec}}
\end{frame}

\section{Experiments}
\begin{frame}
	\frametitle{Noise}
\end{frame}

\begin{frame}
	\frametitle{Discovery}
\end{frame}

\begin{frame}
	\frametitle{Scalability}
\end{frame}

\section{Internet of Things}
\begin{frame}
	\frametitle{Challenges}
	\begin{itemize}
		\item log-comparison method. Performance dependent on model complexity. How many traces are needed to describe the model sufficiently well.
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Benefits}
	\begin{itemize}
		\item Better performance, and especially better scaling w.r.t. number of activities and trace length.
		\item Embeddings have other advantages, no more binary activity equality, could allow for more fine grained conformance.
	\end{itemize}
\end{frame}

\end{document}