### 2021-12-22 slides.tex
Updated slides

\documentclass{beamer}
%Information to be included in the title page:
\title{Conformance Checking Using Embeddings and its Applicability in the Internet of Things}
\author{Jan Kruska}
\date{\today}

\begin{document}

\frame{\titlepage}

\begin{frame}
	\frametitle{Table of Contents}
	\tableofcontents
\end{frame}
\section{Conformance Checking}
\begin{frame}
	\frametitle{Conformance Checking}
\end{frame}

\section{Embeddings}
\begin{frame}
	\frametitle{Embeddings}
\end{frame}

\begin{frame}
	\frametitle{\emph{act2vec} \& \emph{trace2vec}}
\end{frame}

\section{Experiments}
\begin{frame}
	\frametitle{Setup}
	Random trees were generated using pm4py and the following configurations:
	\begin{tabular}
		Min & Mode & Max \\
		5 & 10 & 15 \\
		10 & 20 & 30 \\
		15 & 30 & 45 \\
	\end{tabular}
\begin{tabular}
	Sequence & Parallel & Choice & Loop \\
	0.75 & 0.25 & 0 & 0 \\
	0.75 & 0 & 0.25 & 0 \\
	0.5 & 0.25 & 0.25 & 0 \\
	0.25 & 0.25 & 0.25 & 0.25 \\
\end{tabular}
By combining these parameters 12 trees were generated.
\end{frame}
\begin{frame}
	\frametitle{Example tree}
	\begin{figure}
		\includegraphics[width=1\textwidth]{figures/process-tree}
		\caption{A process tree generated by pm4py using parameters: Min 15, Mode 30, Max 45, sequence 0.75, parallel 0.25. }
		\label{fig:process-tree}
	\end{figure}
\end{frame}

\begin{frame}
\frametitle{Noise}
\end{frame}

\begin{frame}
	\frametitle{Discovery}
\end{frame}

\begin{frame}
	\frametitle{Scalability}
\end{frame}

\section{Internet of Things}
\begin{frame}
	\frametitle{Challenges}
	\begin{itemize}
		\item log-comparison method. Performance dependent on model complexity. How many traces are needed to describe the model sufficiently well.
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Benefits}
	\begin{itemize}
		\item Better performance, and especially better scaling w.r.t. number of activities and trace length.
		\item Embeddings have other advantages, no more binary activity equality, could allow for more fine grained conformance.
	\end{itemize}
\end{frame}

\end{document}