### 2022-01-01 slides.tex
Motivation and structure

\documentclass{beamer}
\usepackage{caption}
\usepackage{subcaption}

\title{Conformance Checking Using Embeddings and its Applicability in the Internet of Things}
\author{Jan Kruska}
\date{\today}

\begin{document}
	\beamertemplatenavigationsymbolsempty
	\addtobeamertemplate{navigation symbols}{}{%
		\usebeamerfont{footline}%
		\usebeamercolor[fg]{footline}%
		\hspace{1em}%
		\insertframenumber
	}
	\frame{\titlepage}

	\begin{frame}
		\frametitle{Table of Contents}
		\tableofcontents
	\end{frame}
	\section{Motivation}
	\begin{frame}
		\frametitle{Motivation}
		\begin{itemize}
			\item We know that traditional approaches (replay and alignments) are able to measure conformance, but have some meaningful disadvantages as well.
			\item At least conceptually NLP and BPM share some important characteristics. Both deal with sequences (sentences/traces) of words/activities and aim to define similarity measures between those.
		\end{itemize}
		\begin{block}{Research Question:}
			Are adaptations of NLP techniques feasible solutions for problems in the BPM field, specifically for the problem of conformance checking.
		\end{block}
		\alert{Theoretical exploration in previous paper, here implementation and evaluation was the main concern}
	\end{frame}

	\section{Structure}
	\begin{frame}
		\frametitle{Structure}
		\begin{figure}
			\includegraphics[width=1\textwidth]{figures/structure}
			\caption{The general structure of the approach}
			\label{fig:structure}
		\end{figure}
	\end{frame}

	\begin{frame}
		\frametitle{Used dissimilarity functions}
		\begin{enumerate}
			\item \textbf{Word Mover's distance} How much "work" is needed to transform one distribution of dirt on piles into another. Both distance between piles and amount of dirt on piles matter.
			\item \textbf{Iterative Constraint Transfers} WMD has cubic complexity in the dictionary size, which makes it unattractive. ICT is a linear time approximation of WMD obtained by relaxing one of the constraints of the LP.
			\item \textbf{\emph{trace2vec}} Analogously to \emph{doc2vec} train an embedding of bags of activities called \emph{trace2vec}. Use cosine similarity to directly compare traces in the trace embedding space.
		\end{enumerate}
	\end{frame}

	\section{Experiments}
	\begin{frame}
		\frametitle{Setup}
		Random trees were generated using pm4py and the following configurations:

		\begin{tabular}{lll}
			Min & Mode & Max \\
			5 & 10 & 15 \\
			10 & 20 & 30 \\
			15 & 30 & 45 \\
		\end{tabular}
		\begin{tabular}{llll}
			Sequence & Parallel & Choice & Loop \\
			0.75 & 0.25 & 0 & 0 \\
			0.75 & 0 & 0.25 & 0 \\
			0.5 & 0.25 & 0.25 & 0 \\
			0.25 & 0.25 & 0.25 & 0.25 \\
		\end{tabular}

		Using these parameters $3\times4=12$ trees were generated.
	\end{frame}
	\begin{frame}
		\frametitle{Example tree}
		\begin{figure}
			\includegraphics[width=1\textwidth]{figures/process-tree}
			\caption{A process tree generated by pm4py using parameters: Min 15, Mode 30, Max 45, sequence 0.75, parallel 0.25. }
			\label{fig:process-tree}
		\end{figure}
	\end{frame}

	\begin{frame}
		\frametitle{Noise}
		\begin{figure}
			\centering
			\begin{subfigure}[b]{0.49\textwidth}
				\centering
				\includegraphics[width=\textwidth]{figures/noise-first}
				\caption{Average of first noise experiment (Varying trace percentage).}
				\label{fig:noise-first}
			\end{subfigure}
			\hfill
			\begin{subfigure}[b]{0.49\textwidth}
				\centering
				\includegraphics[width=\textwidth]{figures/noise-second}
				\caption{Average of second noise experiment (Varying amount of noise applied to trace).}
				\label{fig:noise-second}
			\end{subfigure}
		\end{figure}
	\end{frame}

	\begin{frame}
		\frametitle{Discovery}
		\begin{figure}
			\centering
			\begin{subfigure}[b]{0.49\textwidth}
				\centering
				\includegraphics[width=\textwidth]{figures/fitness}
				\caption{Fitness}
				\label{fig:fitness}
			\end{subfigure}
			\hfill
			\begin{subfigure}[b]{0.49\textwidth}
				\centering
				\includegraphics[width=\textwidth]{figures/precision}
				\caption{Precision}
				\label{fig:precision}
			\end{subfigure}
			\caption{Fitness and Precision for different tree, discovery technique pairings for the three proposed techniques as well as the two verification techniques.}
			\label{fig:discovery}
		\end{figure}
	\end{frame}

	\begin{frame}
		\frametitle{Discovery}
		\begin{itemize}
			\item Measures for proposed methods are not proper fitness metrics, so careful with quantitative analyses!
			\item 
		\end{itemize}
	\end{frame}

	\begin{frame}
		\frametitle{Scalability}
		\begin{figure}
			\includegraphics[width=1\textwidth]{figures/scaling}
			\caption{Runtimes for all three methods for varying sizes of logs and dictionaries.}
			\label{fig:scalability}
		\end{figure}
	\end{frame}

	\section{Internet of Things}
	\begin{frame}
		\frametitle{Challenges}
		\begin{itemize}
			\item log-comparison method. Performance dependent on model complexity. How many traces are needed to describe the model sufficiently well.
		\end{itemize}
	\end{frame}

	\begin{frame}
		\frametitle{Benefits}
		\begin{itemize}
			\item Better performance, and especially better scaling w.r.t. number of activities and trace length.
			\item Embeddings have other advantages, no more binary activity equality, could allow for more fine grained conformance.
		\end{itemize}
	\end{frame}

\end{document}